---
title: Generating Priors of metabolic index parameters using a taxonomic hierarchical
  model
output: pdf_document
date: "2023-07-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(MASS)
library(TMB)
library(ggplot2)
library(gridExtra)
library(readxl)
library(Matrix)
library(mvtnorm)

```


# Background and Motivation

Several recent papers have used compilations of laboratory experiments to estimate parameters that govern the metabolic index: the ratio of oxygen supply to demand. Because taxonomic coverage of these experiments is limited, existing species distributions are also commonly used to make inferences about the critical values of this index below which organisms avoid habitatis, and to predict changes in physiologically-favorable habitat.  

Generally, information sharing between these two approaches have been ad hoc i.e., using the laboratory data to set reasonable bounds on unknown parameters, or using mean values of parameters.  A formal statistical approach (for example, an explicit Bayesian analysis of species distributions and how they are govern by temperature and oxygen) would permit this information sharing to be more explicit and adaptive as more information is collected.

For that reason, an analysis of laboratory data that can provide prior probabilities for in situ observations could enhance our ability to assess how changing temperature and oxygen will change species distributions and ranges.

# Metabolic index

The metabolic index depicts the ratio of oxygen supply to oxygen demand, assuming that both are goverrned by an Arrhenius equation.  A common form of the metabolic index is:

$$ \phi = pO_2 A_o \bigg(\frac{B}{B_{ref}}\bigg)^n \exp\bigg(\frac{E_o}{k_b} \big(\frac{1}{T} - \frac{1}{T_{ref} }\big)\bigg) $$

where $\phi$ is the metabolic index, $pO_2$ is the partial pressure of oxygen (kP), $B$ is body size, $B_{ref}$ is a reference body size,  $k_b$ is Boltzmann's constant, $T$ is temperature (K) and $T_ref$ is a reference temperature, and $A_o$, $n$, and $E_o$ are the parameters to be estimated.  

The parameters of the metabolic index can be estimated from laboratory studies that expose organisms to combinations of temperature and $pO_2$ and estimate the oxygen level at which organisms cannot survive or otherwise exhibit stress that are indicative of insufficient oxygen supply.  By setting these empirically-derived thresholds as the values of $pO_2$ and $T$ that make $\phi$ equal 1, the expression can be solved through ordinary least squared regression on log-transformed rearrangement of the metabolic index.  

Because all observations and experiments are subject to random error, fitting each experimental fitting independently will propagate these errors into the metabolic index parameter estimates. One way to lessen these effects is through hierarchical modeling.  Hierarchical models presume that each parameter for each species follows some probability distribution with a mean and variance set at a higher taxonomic level.  

Recently, Thorson developed a taxonomic hierarchical model to predict life history traits in the absence of independent measurements.  Briefly,  a vector of $j$ traits among organisms at taxonomic level $g$, $\textbf{x}_{g}$ follows a multivariate normal distribution.  The model assumes that these traits evolve, and that taxonomic levels are a proxy for evolutionary potential.  Thus, given a "parent" taxonomic group (the taxonomic group above the group under consideration), the vector $x_{g}$ is:

$$ \textbf{x}_{g} \sim \text{MVN} (\textbf{x}_{[p]}, \Sigma_{l(g)} ) $$
where $\textbf{x}_{[p]}$ are the mean values of the parent taxonomic grouping, and $\Sigma_{l(g)}$ is the variance - covariance matrix for taxonomic level $g$.  The latter is calculated as

$$ \Sigma_{l(g)} = \lambda_{l(g)} \Sigma $$

where $\Sigma$ is the variance-covariance matrices among the parameters, and $\lambda_{l(g)}$ expresses how the variability grows or expands at each taxonomic level. Thorson describes the latter as the proportion of evolutionary covariance occuring at taxonomic level $l(g)$.

# Taxonomic Summary of the data

```{r funs, include = F}
find_index <- function(x,y) y <- which(y == x)
make_df_plot <- function(level, beta_mle, beta_se, ParentChild_gz, groups) {
  group_index <- which(ParentChild_gz$ChildTaxon==level)
  groupname <- groups[level]
  GroupNames <- gsub(".*_","",ParentChild_gz$ChildName[group_index])
  
  
  Est <- tibble("{groupname}" := GroupNames,
                logAo = beta_mle[group_index,1],
                logAomin = beta_mle[group_index,1] - beta_se[group_index,1],
                logAomax = beta_mle[group_index,1]  + beta_se[group_index,1],
                Eo = beta_mle[group_index,3],
                Eomin = beta_mle[group_index,3] - beta_se[group_index,3],
                Eomax = beta_mle[group_index,3] + beta_se[group_index,3],
                n = beta_mle[group_index,2],
                nmin = beta_mle[group_index,2] - beta_se[group_index,2],
                nmax = beta_mle[group_index,2] + beta_se[group_index,2],
  )
  return(Est)
}
plotest <- function(dataest, trait, groupname, xmin, xmax) {
  trait <- enquo(trait)
  groupname <- enquo(groupname)
  xmin <- enquo(xmin)
  xmax <- enquo(xmax)
  
  
  groupplot <- ggplot(data = dataest, aes(x = !!trait, y = !!groupname)) +
    geom_point() + 
    scale_y_discrete(limits = rev) +
    geom_errorbar(aes(y = !!groupname,
                      xmin = !!xmin,
                      xmax = !!xmax)
    ) 
  
  return(groupplot)
}

  rmvnorm_prec <- function(mu, prec, n.sims, random_seed ) {
  set.seed( random_seed )
  z = matrix(rnorm(length(mu) * n.sims), ncol=n.sims)
  L = Matrix::Cholesky(prec, super=TRUE)
  z = Matrix::solve(L, z, system = "Lt") ## z = Lt^-1 %*% z
  z = Matrix::solve(L, z, system = "Pt") ## z = Pt    %*% z
  z = as.matrix(z)
  return(mu + z)
}
sim_spc_in_group <- function(obj, 
                             rep,
                             n.sims, 
                             ParentChild_gz, 
                             Group, 
                             GroupEst, 
                             level, 
                             log_lambda,
                             sigma) {
  
  sim_beta_group <- matrix(NA, nrow = n.sims, ncol = 3)
  sim_beta_species <- matrix(NA, nrow = n.sims, ncol = 3)
  group_index <- which(GroupEst[1] == Group)
  
  # get random draws for all random variables
  newpar = rmvnorm_prec( mu = obj$env$last.par.best,
                         prec = rep$jointPrecision, 
                         n.sims = n.sims,
                         random_seed = sample(1:1000000, 1))
  
  
  parnames <- names(obj$env$last.par.best)
  
  # extract the simulated beta_gj
  beta_gj_random <-newpar[grep(parnames, pattern = "beta_gj"),]
  # assign beta_gj to Ao, no, and Eo matrixes
  class_index <- which(ParentChild_gz$ChildTaxon == 1)
  order_index <-  which(ParentChild_gz$ChildTaxon == 2)
  family_index <-  which(ParentChild_gz$ChildTaxon == 3)
  species_index <- which(ParentChild_gz$ChildTaxon == 4)
  n_gcj <- length(class_index)
  n_goj <- length(order_index)
  n_gfj <- length(family_index)
  n_gj <- length(species_index)
  
  # make an array for looking things up
  lookup_array <- c(rep("Ao_class", n_gcj),
                    rep("Ao_order", n_goj),
                    rep("Ao_family", n_gfj),
                    rep("Ao_species", n_gj),
                    rep("n_class", n_gcj),
                    rep("n_order", n_goj),
                    rep("n_family", n_gfj),
                    rep("n_species", n_gj),
                    rep("Eo_class", n_gcj),
                    rep("Eo_order", n_goj),
                    rep("Eo_family", n_gfj),
                    rep("Eo_species", n_gj)
  )
  
  
  allgroups <- c("class", "order", "family")
  # Put all estimates for the corresponding taxonomic level into a matrix, 
  # each row is a group at that level, columns are random draws
  
  Ao_sim <- beta_gj_random[which(lookup_array == paste0("Ao_", allgroups[level])), ]
  no_sim <- beta_gj_random[which(lookup_array == paste0("n_", allgroups[level])), ]
  Eo_sim <- beta_gj_random[which(lookup_array == paste0("Eo_", allgroups[level])), ]
  
  # save results to sim_beta_class by looking up corresponding row
  
  sim_beta_group[,1] <- c(Ao_sim[group_index,])
  sim_beta_group[,2] <- c(no_sim[group_index,])
  sim_beta_group[,3] <- c(Eo_sim[group_index,])
  
  # using random draw for Class, simulate mean for a random order, random family, and then the value for a given species
  if (level ==1) {
    sim_beta_ord <- sim_beta_group + rmvnorm(n = n.sims, mean = rep(0,3), sigma = exp(log_lambda[1]) * sigma)}
  if (level == 2) sim_beta_ord <- sim_beta_group
  if (level <=2) sim_beta_fam <- sim_beta_ord + rmvnorm(n = n.sims, mean = rep(0,3), sigma = exp(log_lambda[2]) * sigma)
  if (level ==3) sim_beta_fam <- sim_beta_group
  sim_beta_species <- sim_beta_fam + rmvnorm(n = n.sims, mean = rep(0,3), sigma = exp(log_lambda[3]) * sigma)
  # Save the species result into matrix
  #sim_beta_species[i,] <- sim_beta_spc
  #sim_beta_species[i,] <- c(Ao_sim[actinop_index], no_sim[actinop_index], Eo_sim[actinop_index])
  #}
  sim_beta_df <- tibble(logAo = sim_beta_species[,1],
                        Eo = sim_beta_species[,3],
                        n = sim_beta_species[,2],
                        Group = Group)
  
  
  return(sim_beta_df)
}
```


The following tables depict the data organized by Class, Order, and Family.
```{r loadsummarise, include = F} 
### Generate Evolutionary Trait Structure ####

#### Get taxonomy tree ####
all.dat <- readRDS(file = "data/alldata_taxonomy.RDS")
# remove data where there is no body size data
all.dat <- dplyr::filter(all.dat, !is.na(W))
naIndex <- which(is.na(all.dat$Species))
for (i in 1:length(naIndex)) all.dat$Species[naIndex[i]] <- paste0(all.dat$Genera[naIndex[i]], " spc")

# get median mass for each species
species.median.mass <- all.dat %>%
  group_by(Species) %>%
  summarize(Wmed = median(W))
# divide Actual mass by median mass for that species
for (i in 1:nrow(species.median.mass)) {
  spc.index <- which(all.dat$Species == species.median.mass$Species[i])
  all.dat$W[spc.index] <- all.dat$W[spc.index] / species.median.mass$Wmed[i]
}
# for species with only 1 mass, replace the NA with 1
all.dat$W[is.na(all.dat$W)] = 1
# describe the data a bit - how many unique  order per class, families per order, etc.
class_summary <- all.dat %>%
  group_by(Class) %>%
  summarise(NoOrder = length(unique(Order)), NoFamily = length(unique(Family)),  NoSpecies = length(unique(Species)))
print(class_summary, n = 30)

order_summary <- all.dat %>%
  group_by(Order, Class) %>%
  summarise(NoFamily = length(unique(Family)), NoSpecies = length(unique(Species)))
print(order_summary, n = 50)

family_summary <- all.dat %>%
  group_by(Family, Order) %>%
  summarise(NoGenus = length(unique(Genera)), NoSpecies= length(unique(Species)))
print(family_summary, n = 70)

rem.classes <- FALSE
# Remove class with a single order
if (rem.classes) {
keep.class <- class_summary$Class[class_summary$NoOrder>1]
all.dat <- dplyr::filter(all.dat, Class %in% keep.class)
}
```

```{r maketables, echo = F} 
knitr::kable(class_summary, caption = "Summary of data by Class")
knitr::kable(order_summary, caption = "Summary of data by Order")
knitr::kable(family_summary, caption = "Summary of data by Family")
```
By far, the most well represented groups are the ray-finned fishes (Class Actinopteri) and the Malacostraca crustaceas (Classs Malacostraca).  Of the latter, roughly half were in the Order Decapoda.  For the vast majority of Orders, the dataset included only a single Family.  Within families, there were at most two genera and two species represented.  

Because of the limited number of organisms within lower taxonomic groupings in the database, we ran the hierchircal model using four taxonomic groupings Class, Order, Family and Species (i.e, we omitted Genus)

# Hierarchical Model Findings Summary

### Point 1: n does not vary, and is approximately -0.1.

At nearly every taxonomic level (Class, Order, Family), the mean value of n did not vary substantially from the grand mean of -0.08.  A small number of Family-level estimates showed some deviation, but these were all cases in which a single species and experimental study was represented.

```{r fitmodel, echo = F, include = F}
### Setup TMB data and parameters ####
#### Create new ParentChild matrix for reduced taxonomic structure ####
kb <-  8.617333262145E-5
tref <- 15
all.dat$inv.temp <- (1 / kb) * (1 / (all.dat$Temp + 273.15) - 1/(tref + 273.15))
all.dat$Pcrit_atm <- all.dat$Pcrit / 101.325 # convert from KPa to atm
all.dat$minuslogpo2 <- - log(all.dat$Pcrit_atm) 
taxa.list <- c("Class", "Order", "Family", "Species")


Z_ik_main <- dplyr::select(all.dat, all_of(taxa.list))
Z_ik <- unique(Z_ik_main, MARGIN = 1)
ParentChild_gz = NULL
# 1st column: child taxon name
# 2nd column: parent taxon name
# 3rd column: parent row-number in ParentChild_gz
# 4th column: Taxon level
# Loop through
for( colI in 1:ncol(Z_ik)){
  Taxa_Names = apply( Z_ik[,1:colI,drop=FALSE], MARGIN=1, FUN=paste, collapse="_")
  Unique_Taxa = unique(Taxa_Names)
  for( uniqueI in 1:length(Unique_Taxa) ){
    Which = which( Taxa_Names == Unique_Taxa[uniqueI] )
    if( colI==1 ){
      ParentChild_gz = rbind( ParentChild_gz, c(Unique_Taxa[uniqueI], NA, NA, colI) )
    }else{
      if( length(unique(Z_ik[Which,colI-1]))>1 ) stop("Taxa has multiple parents")
      ChildName = Unique_Taxa[uniqueI]
      ParentName = paste(rev(rev(strsplit(ChildName,"_")[[1]])[-1]),collapse="_")
      ParentChild_gz = rbind( ParentChild_gz, c(ChildName, ParentName, match(ParentName,ParentChild_gz[,1]), colI) )
    }
  }
}

# Relabel
ParentChild_gz = data.frame( ParentChild_gz )
colnames(ParentChild_gz) = c("ChildName", "ParentName", "ParentRowNumber", "ChildTaxon")
ParentChild_gz[,'ParentRowNumber'] = as.numeric(as.character(ParentChild_gz[,'ParentRowNumber']))
ParentChild_gz[,'ChildTaxon'] = as.numeric(as.character(ParentChild_gz[,'ChildTaxon']))
PC_gz<- as.matrix(ParentChild_gz[, c('ParentRowNumber', 'ChildTaxon')]) - 1
# Identify location for every observation
Taxa_Names = apply( Z_ik, MARGIN=1, FUN=paste, collapse="_")
g_i = match( Taxa_Names, ParentChild_gz[,'ChildName'] )
n_k = ncol(Z_ik)
n_j = 3 # three traits
n_g = nrow(ParentChild_gz)
n_i <- length(g_i)

#### Create index of data to Parent - Child ####
#Z_ik_dat <- dplyr::select(all.dat, Class, Order, Family, Species)
Taxa_Names_dat <-  apply( Z_ik_main, MARGIN=1, FUN=paste, collapse="_")
g_i_dat = match( Taxa_Names_dat, ParentChild_gz[,'ChildName'] )
g_i_i <- sapply(FUN = find_index, X = g_i_dat, y = g_i)

# Create index of species to Parent  - Child
spc_in_PC_gz <- which(PC_gz[,2] == max(PC_gz[,2]))

# Setup TMB ####
data <- list(PC_gz = PC_gz,
             g_i = g_i - 1,
             invtemp = all.dat$inv.temp,
             logW = log(all.dat$W),
             taxa_id = g_i_i -1,
             minuslogpo2 = all.dat$minuslogpo2,
             spc_in_PCgz = spc_in_PC_gz -1
             
)

parameters = list(alpha_j = rep(0,n_j),
                  L_z = rep(1, 6),
                  log_lambda = rep(0, length(unique(PC_gz[,2])) -1),
                  beta_gj = matrix(0, nrow = n_g, ncol = n_j),
                  logsigma = 0
)
Random <- c("beta_gj")

model <- "hierarchical_mi"
compile(paste0("code/TMB/", model, ".cpp"))
dyn.load(dynlib(paste0("code/TMB/",model)))


obj <-
  MakeADFun(
    data = data,
    parameters = parameters,
    DLL = model,
    random = Random,
    silent = TRUE
  )
opt <- nlminb(obj$par, obj$fn, obj$gr)
rep = sdreport( obj,
                getReportCovariance = TRUE, 
                getJointPrecision=TRUE)
summary(rep, "fixed")
re <- summary(rep, "report")


spc_parameters <- re[grep(rownames(re), pattern  = "spc_ij"),1]
spc_parameters_se <-re[grep(rownames(re), pattern  = "spc_ij"),2]
spc_ij_mle <- matrix(spc_parameters, nrow = n_i, ncol = 3, byrow = F)
spc_ij_se <- matrix(spc_parameters_se, nrow = n_i, ncol = 3, byrow = F)


re <- summary(rep, "random")
fixef <- summary(rep, "fixed")
beta_mle <- matrix(re[grep(rownames(re), pattern = "beta"),1], nrow = n_g, ncol = 3, byrow = F)
beta_se <- matrix(re[grep(rownames(re), pattern = "beta"),2], nrow = n_g, ncol = 3, byrow = F)

```

```{r show_n, echo = F, fig.cap="Estimated mean (SE) of n for Class, Order, and Family", fig.height = 7}
ClassEst <- make_df_plot(level = 1, 
                         beta_mle,
                         beta_se,
                         ParentChild_gz,
                         groups = taxa.list)

Est.2.plot <- merge(ClassEst, class_summary)
Est.2.plot <- ClassEst
Aoplot <- plotest(Est.2.plot, logAo, Class, logAomin, logAomax)
Eoplot <- plotest(Est.2.plot, Eo, Class, Eomin, Eomax)
nplotclass <- plotest(Est.2.plot, n, Class, nmin, nmax)


#### Plot Orders #####

OrderEst <- make_df_plot(level = 2, 
                         beta_mle,
                         beta_se,
                         ParentChild_gz,
                         groups = taxa.list)

OrderEst <- merge(OrderEst, order_summary)
Est.2.plot <- OrderEst
Aoplot <- plotest(Est.2.plot, logAo, Order, logAomin, logAomax)
Eoplot <- plotest(Est.2.plot, Eo, Order, Eomin, Eomax)
nplotorder <- plotest(Est.2.plot, n, Order, nmin, nmax)



#### Plot Families ####
FamilyEst <- make_df_plot(level = 3, 
                         beta_mle,
                         beta_se,
                         ParentChild_gz,
                         groups = taxa.list)
FamilyEst <- merge(FamilyEst, family_summary)
Est.2.plot <- FamilyEst
Aoplot <- plotest(Est.2.plot, logAo, Family, logAomin, logAomax)
Eoplot <- plotest(Est.2.plot, Eo, Family, Eomin, Eomax)
nplotfamily <- plotest(Est.2.plot, n, Family, nmin, nmax)

grid.arrange(nplotclass, nplotorder, nplotfamily,
             layout_matrix = rbind(c(1, 3),
                        c(2, 3)))

```

## Point 2: Few Differences in Ao and Eo among classes and orders

Below I show results of simulating 10,000 random species within each taxomic grouping, where the group-level means were drawn from multivariate normal disributions, and then then the estimated lower-level values down to species were simulated based on the estimated mean values,estimated variance covariance matrix and the $\lambda_{l(j)}$.  These would be similar to a prior probabilty if one knew the taxonomic identity of an organism at a high level (Class, Order, or Family) and wished assign a prior probability at the species level.  

```{r getsigma, echo = F} 
### Make Sigma ####
#### Make empty Cholesky Matrix ####
L <- matrix(0, nrow = n_j, ncol = n_j)
# extract L_z
fixed_names <- rownames(fixef)

L_z <- fixef[grep("L_z", fixed_names),1]
#### Fill iin Cholesky Matrix ####
Count = 1
for (i in 1 : n_j) {
  for (j in 1 : n_j) {
    if (i>=j) {
      L[i,j] <- L_z[Count]
      Count <- Count + 1
    }
  }
}

##### Calculate Sigma and extract log_lambdas
sigma <- L %*% t(L)
log_lambda <- fixef[grep("log_lambda",fixed_names),]
```

```{r plotclass, echo = F, fig.cap = "2-D Density of Ao and Eo for species drawn within each Class" }
#### Plot All Classess ####

sim_beta_df <- NULL
GroupEst <- ClassEst
ngroups <- nrow(GroupEst)
n.sims <- 10000
for (i in 1:ngroups) {
  Group.2.use <- GroupEst$Class[i]
  sim_beta_df_group <- sim_spc_in_group(obj  = obj,
                                        rep = rep,
                                        n.sims = n.sims,
                                        ParentChild_gz = ParentChild_gz,
                                        Group = Group.2.use ,
                                        GroupEst = GroupEst,
                                        level = 1,
                                        log_lambda = log_lambda, 
                                        sigma
  )
  sim_beta_df <- rbind(sim_beta_df, sim_beta_df_group)
}


d <- ggplot(data = sim_beta_df, aes( x = logAo, y = Eo)) + 
  geom_density_2d_filled( stat = "density_2d_filled", h = c(1, 0.2),
                          show.legend = F) +
  xlim(1.5, 5.5) + 
  ylim(0, 0.8) + 
  facet_wrap(vars(Group), nrow = 4, ncol = 3)
print(d)  

```
Generally, the middle of the simulated species within Classes are close to the overall mean values (Ao = 3.41, Eo = 0.350), and 95% of species in these classes will fall between 1.95 and 4.88 (Ao) and 0.036 and 0.66 (Eo).

There were slight differences among Orders, where only Orders with at least 3 separate species are shown.


```{r plotorders, echo = F, fig.cap = "Same plot for Orders that had at least three species represented in the data"}

sim_beta_df <- NULL
redOrderEst <- dplyr::filter(OrderEst, NoSpecies >=3)
ngroups <- nrow(redOrderEst)
for (i in 1:ngroups) {
  Group.2.use <- redOrderEst$Order[i]
  sim_beta_df_group <- sim_spc_in_group(obj  = obj,
                                        rep = rep,
                                  n.sims = n.sims,
                                  ParentChild_gz = ParentChild_gz,
                                  Group = Group.2.use ,
                                  GroupEst = redOrderEst,
                                  level = 2,
                                  log_lambda = log_lambda, 
                                  sigma
                                  )
  sim_beta_df <- rbind(sim_beta_df, sim_beta_df_group)
}
  

d <- ggplot(data = sim_beta_df, aes( x = logAo, y = Eo)) + 
  geom_density_2d_filled( stat = "density_2d_filled", h = c(1, 0.2),
                          show.legend = F) +
  xlim(1.5, 5.5) + 
  ylim(0, 0.8) + 
  facet_wrap(vars(Group), nrow = 2, ncol = 2)
print(d)  

```

Slightly larger differences were projected among some families, although many of these groups included only two species (three families had three species represented). Zoocaridae (eelpouts) and Gadidai (cods) had relatively low Ao (or lower tolerance for low $pO2$), while species in the families Penaidae and Triptyrygiidae had a higher Ao.  There was more variability in $E_o$ among families, with estimated temperature dependence being low among Salmonidae and exceptionally high among Penaeidae.

```{r plotfam, fig.cap = "Same as above but for Family, restricted to Families represented by at least two species"}
#### Plot families with at least 2 species ####
sim_beta_df <- NULL
redFamilyEst <- dplyr::filter(FamilyEst, NoSpecies >=2)
ngroups <- nrow(redFamilyEst)
for (i in 1:ngroups) {
  Group.2.use <- redFamilyEst$Family[i]
  sim_beta_df_group <- sim_spc_in_group(obj  = obj,
                                        rep = rep,
                                        n.sims = n.sims,
                                        ParentChild_gz = ParentChild_gz,
                                        Group = Group.2.use ,
                                        GroupEst = redFamilyEst,
                                        level = 3,
                                        log_lambda = log_lambda, 
                                        sigma
  )
  sim_beta_df <- rbind(sim_beta_df, sim_beta_df_group)
}


d <- ggplot(data = sim_beta_df, aes( x = logAo, y = Eo)) + 
  geom_density_2d_filled( stat = "density_2d_filled", h = c(1, 0.2),
                          show.legend = F) +
  xlim(1.5, 5.5) + 
  ylim(-0.1, 1.0) + 
  facet_wrap(vars(Group), nrow = 4, ncol = 4)
print(d)  
```
# Point 3: Hierarchical model improves estimation through shrinkage

Fitting individual linear model to species-level data typically leads to wider variability of traits than hierarchical models. In other words, the hierarchical model dismisses exceptionally unusual parameter estimates to a species data and "shrinks" those estimates to the Parent group level means.  In essence, this stabilizes parameter estimates.

To do: fit each species independently when there are adequate T, W data

Note: Cannot recreate parameter estimates in spreadsheet "MI_traits_Est_by_Curtis"

